---
title: "Assigment 1"
author: "Fabi??n Garc??a"
date: "July 17, 2014"
output: html_document
---

# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

1. Setting working directory
```{r}
setwd('/Users/fabian/Dropbox/Coursera/RepData_PeerAssessment1')
````

2. Unzipping data
```{r}
unzip('activity.zip')
```

3. Loading data
```{r}
activity <- read.csv('activity.csv')
```

4. Checking activity content (first 5 rows)
```{r}
head(activity)
```

5. Changing the class of date column
```{r}
activity$date <- as.Date(activity$date)
```

6. New dataframe summing steps over each day
```{r}
dayly_data <- aggregate(formula = steps ~ date, data = activity, FUN = sum, na.rm = T)
```

6. Histogram of the total number of steps taken each day
```{r}
hist(dayly_data$steps, main="Mean of steps per day", xlab="# of steps")
```

## What is mean total number of steps taken per day?
6. Mean of steps per day
```{r}
mean(dayly_data$steps, na.rm=TRUE)
```
7. Median of steps per day
```{r}
median(dayly_data$steps, na.rm=TRUE)
```
## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
